Colonies - Compute Continuums across Platforms

### 2023-04-08 paper_2023/paper.tex
wip

The Colonies framework uses a combination of technologies to address the aforementioned challenges. A fundamental design principle is stateless web serices, which makes Colonies simpler, more reliable and scalable, and easier to implement. To ensure reliability and data consistency, distributed consensus algorithms are needed to offer high-availability and handle server crashes. The remainder of this section describes how the Colonies framework is implemented. The next discusses the role of queues to support batch jobs and realtime processing, but also how queues can be used to enable loose coupling of system components.

@@

\subsection{A stateless failsafe mechanism}
The primary objective of all Colonies API requests is to alter some state stored in the database or retrieve information from the database. The Colonies framework is designed to be stateless, meaning that the Colonies server does not keep any information between requests. In other words, each request is handled independently, without relying on any information from previous requests.

@@

\begin{figure}[h]
	\centering
    \includegraphics[scale=0.5]{raft.png}
	\caption{High-availability deployment of Colonies servcer.}
	\label{fig:ha_deployment}
\end{figure}

@@

\subsubsection{Cron}
TODO
\begin{figure}[h]
	\centering
    \includegraphics[scale=0.4]{cron.png}
	\caption{Sample figure caption.}
	\label{fig:fig1}
\end{figure}

\subsubsection{Generators}
TODO

\subsubsection{Zero-trust security}
\label{zerotrustsecurity}
TODO

>>>>

The Colonies framework uses a combination of technologies to address the aforementioned challenges. A fundamental design principle is statelessness, which makes Colonies simpler, more reliable and scalable, and easier to implement. To ensure reliability and data consistency, distributed consensus algorithms are needed to offer high-availability and handle server crashes. The remainder of this section describes how the Colonies framework is implemented. The next discusses the role of queues to support batch jobs and realtime processing, but also how queues can be used to enable loose coupling of system components.

@@

\subsection{A stateless failsafe mechanism}
The primary objective of all Colonies API requests is to alter some state stored in the database or retrieve information from the database. The Colonies framework is designed to be stateless, meaning that the Colonies server does not keep any information in memory between requests. Each request is handled independently without relying on information from previous requests.

@@

\begin{figure}[h]
	\centering
    \includegraphics[scale=0.5]{raft.png}
	\caption{High-availability deployment of Colonies.}
	\label{fig:ha_deployment}
\end{figure}

@@

\subsubsection{Cron}
The Colonies framework is based on a stateless architecture where each request is treated as an independent transaction and is processed without any information from prior requests. Since there is no session data stored in memory, Colonies can easily switch between servers without disruptions or data losses. In a stateful architecture on the other hand, the server must maintain a record of connected clients' state and session data in memory, which can result in complications if a server fails or crashes.

Cron is a time-based job scheduler designed to trigger workflows at pre-defined intervals. It can for example be used to automate tasks such as routinely fetching data from external systems. To make Colonies robust and easy to implement it is essential to make Cron workflows stateless. This can be achieved by introducing an additional database table to store information about Cron workflow, and let the elected Colonies server leader assumes the responsibility of managing Cron workflows through a two-step process.

Firstly, the leading Colonies server calculates a future deadline, indicating when a specific Cron workflow is scheduled to be executed. The leader server then conducts periodic scans of the Cron table. When the current time surpasses the calculated deadline for a particular workflow, a new workflow is automatically submitted. Note that this protocol maintains complete statefullness as no session information is stored in memory between cron scans. In the event that the leader Colonies server experiences a crash, a new leader is simply selected and assumes the responsibility of managing Cron workflows according to the two-step protocol.

\subsubsection{Generators}
Another features of the Colonies framework are so-called generators. Generators are workflow templates that are triggered automatically when specific conditions are met. Generators can for example be used accumulate input data in bulks to improve performance or parallelism. This is achieved by sending a special \emph{pack request} to a Colonies server, targeting a certain generator. Generators facilities integration with third-party systems as it allows for stateless implementation where interactions with Colonies can occur using a fire-and-forget style. Colonies guarantees that all pack request submitted to a generator is evenutally processed, even if a Coloniers server is restarted or fails.

Without generators, third-party software would have to use an alternative method such as a message broker or a database to enqueue input data and periodically empty the queues to generate workflows themselves. This adds complexity to the integration process and may requires additional infrastructure components. Generators are a powerful feature of the Colonies framework that can help simplify integration and implement robust data processing pipelines.

Generators can be implemented using a similar method to how Cron workflow are implemented. The elected leader Colonies server is responsible for scanning a special generator table database to decide if a workflow should be triggered. This approach ensures that there is a single point of control for generating workflows and prevents generation of duplicate workflows.

If the leader server crashes, a new leader Colonies server is automatically elected, and it takes over the responsibility for triggering the generators. This ensures that the generation of workflows is not impacted by server failures, and the system can continue to operate smoothly even in case of failures. It can be worth to point out that since the pack request only adds information to the database and does not manipulate any states, it can be processed by any Colonies server without synchronization precautious. This means that any server in the cluster can handle incoming pack requests, which increases the scalability and fault-tolerance of the system.

\subsubsection{Zero-trust security}
\label{zerotrustsecurity}
Zero-trust security is a fundamental to the Colonies framework.

\begin{table}[h]
	\caption{Colony Id: 4787a5071856a4acf702b2ffcea422e3237a679c681314113d86139461290cf4  }
	\centering
	\begin{tabular}{ll}
		\toprule
		\cmidrule(r){1-2}
        Executor Id & Executor Name \\
		\midrule
        8a491bcac0be623a54411dd0934fdcdc9c844de5700527a7dbc9da08a6a8310d & $E_{1}$ \\
        6a65f40343999415f47e739f09b625ab437b069b5f591f9844c234533f505bee & $E_{2}$ \\
        41e4c1b10f92a53b7a5a86620ed366635901a1a96c48312a0ef566a13217fe03 & $E_{3}$ \\
		\bottomrule
	\end{tabular}
	\label{deptable}
\end{table}
